### 2018-12-12 doc/paper.tex
paper updates

% notes
% \documentclass[format=sigconf, review=true, draft=true, screen=true]{acmart}
\documentclass[english,preprint,JIP]{ipsj}
% \documentclass[sigplan,authorversion,draft]{acmart}

% \usepackage[utf8]{inputenc}

\usepackage{textcomp}
\usepackage{comment}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphics,graphicx}
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{auto-pst-pdf}

% from matias
\usepackage{balance}

\usepackage{listings}
\lstset{
language=ruby,
% backgroundcolor=\color[rgb]{0.95, 0.95, 0.95},
tabsize=2,
rulecolor=,
basicstyle=\ttfamily,
upquote=true,
% aboveskip={1.5\baselineskip},
columns=fullflexible,
% columns=fixed,
showstringspaces=false,
extendedchars=true,
breaklines=true,
prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
frame=shadowbox,
showtabs=false,
showspaces=false,
showstringspaces=false,
% identifierstyle=\ttfamily,
keywordstyle=\color[rgb]{1.0,0,0}\bfseries,
keywordstyle=[1]\color[rgb]{0,0,0.75}\bfseries,
keywordstyle=[2]\color[rgb]{0.5,0.0,0.0},
keywordstyle=[3]\color[rgb]{0.127,0.427,0.514},
keywordstyle=[4]\color[rgb]{0.4,0.4,0.4},
commentstyle=\color[rgb]{0.5,0.5,0.5}\itshape,
stringstyle=\color[rgb]{0.639,0.082,0.082},
morekeywords={read-lines, take, each, produce, consume, put, get, make-channel},
deletekeywords={do, read},
%numbers=left,%
numbersep=5pt,%
numberstyle=\tiny\color{gray},%
emphstyle=\bfseries,%
breaklines=true,
breakatwhitespace=true,%
escapechar=`,
%numbers=left,
}
\lstset{escapeinside={<@}{@>}}
\definecolor{ttblue}{rgb}{0,0,0.75}

\usepackage[varg]{txfonts}%%!!
\makeatletter%
\input{ot1txtt.fd}
\makeatother%

\begin{document}
\title{A Shell-like Model for General Purpose Programming}

\affiliate{TITech}{東京工業大学情報理工学院数理・計算科学系 \\
Department of Mathematical and Computing Science, Tokyo Institute of Technology,
Meguro, Tokyo 152-8552}
\affiliate{KTH}{Kungliga Tekniska H\"ogskolan\\
KTH Royal Institute of Technology, Stockholm, Sweden, 10044}

\author{Jeanine Miller Adkisson}{TITech}[jneen@prg.is.titech.ac.jp]
\author{Hidehiko Masuhara}{TITech}[masuhara@acm.org]
\author{Johannes Westlund}{TITech,KTH}[jwestlun@kth.se]

% thought experiment---parallel fetching
% find an example of named pipes

% TODO LIST
% * Find an example of named pipes reopening
% * Find an example of a non-linear pipe arrangement (prob. from flow-based programming)
% * Plan the Description of Magritte section
% * Add Call Semantics to Design Considerations
% * Institute and email addr from Johannes
% * Explain why we do not do eager interruption (limit sources of interruptions)
% * Finish Non-Interruption section
% * Completely redo the Magritte section
% * Bibliography
% * Pretty Syntax

\begin{abstract}
Shell languages such as bash are designed to integrate with an OS, which mainly involves managing processes with implicit input and output streams. They also attempt to do this in a compact way that could be reasonably typed on a command-line interface.

However, existing shell languages are not sufficient to serve as general-purpose languages---values are not observable except in raw streams of bytes, and they lack modern language features such as lexical scope and higher-order functions.

By way of a new programming language, \emph{Magritte}, we propose a general-purpose programming language with semantics similar to bash. In this paper, we discuss the early design of such a system, in which the primary unit of composition, like bash, is processes with asynchronous inputs and outputs, which can be read from or written to at any time, and which can be chained together via a pipe operator. We also explore concurrency semantics for such a language.

% We explore the implementation and design challenges that come with such a model, and also show that this model is able to easily create and compose concurrent algorithms.
% The UNIX shell programming model has been shown to be a successful model for composing programs---spawning unrelated programs in parallel to communicate over operating-system pipes.

% Beyond the most simple tasks, however, \texttt{bash} and similar tools break down, due to the byte-based nature of the system: variables can only hold byte arrays, and pipes can only manage raw byte streams, with no support for transactional writes.
% %
% We explore a pipe-based programming model that attempts to synthesize the good parts of UNIX shells, departing at key points, along with a supporting concurrency model and a semantic of interruption.
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Functional and object-oriented programming suffer from a disconnect between the concept of \emph{functions} and \emph{programs}.  Whereas functions and methods receive values as arguments and return single values as a result, programs can receive data not only as arguments but also asynchronously through an input, and write results asynchronously to an output.
%
% In functional and object-oriented languages, this can result in not only a significant amount of abstraction overhead for dealing with external functions, but the immediacy of function arguments and return values can make composing concurrent programs difficult.
%
% By way of a new programming language, \emph{Magritte}, we propose a programming model in which the primary unit of composition is not immediate function arguments and return values, but instead processes with asynchronous inputs and outputs, which can be read from or written to at any time, and which can be chained together via a pipe operator, much like programs in a UNIX environment. We also explore a concurrency semantic for such a language.
%
% We explore many implementation and programming challenges that come with such a model, and also show that this model is able to easily create and compose concurrent algorithms that would be difficult to write or compose in a functional or object-oriented language.

% Despite the widespread use of GUI tools for configuration and orchestration tasks, CLI tools, and more specifically shell-based interfaces remain a powerful standard for software distribution and integration. Several attempts have been made at solving the desktop-integration problem: how should we allow programs written in different languages and on different platforms to be scripted and integrated in a coherent system? The most notable attempts at this problem have been TCL(TODO: cite!) and Guile Scheme(TODO: cite!), but neither has been as succesful as they intended to be: the common language for a large portion of tools remains shell-based, relying on such concepts as standard I/O, string-based argument vectors, and the expectation that users will be able to freely pipe streams of bytes in a pipe-based way.
%
% Unfortunately, existing UNIX shell languages have significant known interoperability issues, such as lack of any kind of usable encapsulation, and the lack of in-language data structures to represent any kind of value other than a string. We believe that a shell-compatible language with well-defined semantics that could also function as a general-purpose systems language could go a long way towards solving the desktop-integration problem.
%
% On the other hand, a persistent problem with more established systems like Erlang or Go is that they don't account for closing channels, and require manual cleanup of processes. In a systems programming context, errors and hangups are commonplace and expected, and it is critical to have some measure of automatic process cleanup for even naively-written code.
%
% We propose a channel-based communication system that integrates with a workflow compensation system and is capable of recovering from errors and, importantly, cleaning up processes when communication halts.
\end{abstract}

\maketitle

\section{Motivation}

The UNIX shell programming model has been shown to be a successful model for composing programs---spawning unrelated programs in parallel to communicate over operating-system pipes. Beyond the most simple tasks, however, \verb/bash/ and similar tools break down, due to the byte-based nature of the system: variables can only hold byte arrays, and pipes can only manage raw byte streams, with no support for transactional writes.

In this paper, we use the term \emph{pipe-based language} to mean a programming language intended to be used on a command line, with the ability to spawn many processes in which communicate through synchronous channels or pipes in an ad-hoc manner --- and in which this facility is the primary method of composing different functions or units.

Our overarching goal is to create a language and a programming system that retains the pipe-based programming and interaction model of bash, but allows for large programs in a way that existing shell languages do not.

\section{Design Considerations}
In this section we discuss several feature requirements and design considerations for a pipe-based language to be viable for large programs.

\paragraph{Not Everything Is a String.}
The first core issue that prevents most shells from being viable for large programs is that only strings are allowed as first-class values.

Several projects\cite{haahr}\cite{duff}, including bash itself, have attempted to add lists and hash-maps as standard objects. Unfortunately, while they allow these values to be stored in variables, most projects do not allow these data structures to be passed into or returned from functions, or importantly, passed through pipes normally.
Since values in these languages cannot be passed through pipes, they are relegated to a second class of value, one for which pipe-based composition is not available.

Thus if we are to allow for pipe-based composition to be the core composition method for large programs, we must allow rich values to be passed through channels.

\paragraph{Transactional Channels.}
Since UNIX pipes can only manage bytes, usually channels with a fixed buffer are specified for performance reasons. However, in the languages we consider, we assume we can transactionally write a single complex value to a channel, resulting in the ability to send data in batch without resorting to a queue that may interleave writes.

Consider the fairly common architecture of producer/consumer: A \emph{producer} process produces values that are processed in parallel by multiple \emph{consumer} processes, and then the values are collected in a single output. This might be expressed in bash as:
\begin{lstlisting}
# pass through the stream, labeling every line
label() {
  while read line; do echo "$1$line"; done
}

# process the stream with different labels
split() {
  label a &
  label b &
}

# process the numbers 1-100, limiting the
# output to the first 3 lines
seq 100 | split | head -3
\end{lstlisting}

Unfortunately, when we run this process, the outputs from the \verb/label/ function become interleaved, resulting in outputs such as:

\begin{verbatim}
a1    b12   a
a2    a     b12
a3    a4    a34
\end{verbatim}
In a language intended for large programs, therefore, writes to a channel must be \emph{transactional}---a single write must be sent wholesale to a single reader without any interleaving, no matter how big the write.

Luckily, given the decision to allow rich values to be passed through channels, a buffer is not as necessary for performance purposes, since a single write may contain an arbitrarily large amount of data---or for that matter a pointer.

\paragraph{Return Semantics.}
Pipeline elements such as these do not generally return values---instead, they output data to their output stream. However, for serious programs it is desirable to capture some kind of return value from functions. ES Shell \cite{haahr} accomplishes this by introducing a \emph{return value} which is separate from stream output and can be accessed via special-purpose call syntax.

If we allow values to be written to output streams, however, this approach is redundant---a function author must decide whether to return a value or write it to standard output. Thus it is desirable to have an interpolation syntax which \emph{collects} the output of a command synchronously.

A priori, such a collection mechanism would have to return an array, as any function call may output zero or more values. However, that approach would lead to the callers of functions needing to manually unwrap arrays on every function call. In order to avoid this, we \emph{expand} the return values into the command vector---increasing the argument number by the number of values output from the function.

In Magritte, normal parentheses in argument position will collect and expand output values in exactly this manner:
\begin{lstlisting}
# a function defintion: output three values
(count-three) = (put 1; put 2; put 3)

# a function call, using the above function
other-fn 0 (count-three) 4

# expands to
other-fn 0 1 2 3 4
\end{lstlisting}

In this way, we can use output values in exactly the same way as return values seamlessly. These semantics are similar to the \verb/$(...)/ syntax in bash, with the exception that because we write transactional values to the output, we have the ability to properly separate values without relying on whitespace.

\paragraph{Variable Scope.}
Like the Es shell\cite{haahr}, we also wish to provide lambda functions with lexically scoped variables. However, Es has an ambiguity with respect to its variables---since both lexically scoped variables and dynamically scoped variables share the same syntax, the scope of a variable depends on runtime information. Additionally, since Es has no syntactic difference between let-binding and mutation, even code with no free variables may change in meaning signifcantly depending on what variables are in scope in its context.

To remedy this issue, we introduce a separate syntax for lexically scoped variables, using a \verb/%/ instead of \verb/$/ (for example, \verb/%x/). Let-binding will bind a variable both lexically and dynamically, and uses an unadorned variable (e.g. \verb/x = 1/), so that we can maintain compatibility with standard environment files. Variable binders in lambda arguments are similarly both lexical and dynamic, and use \verb/?/ (e.g. \verb/?x/) to allow for a planned extension for pattern matching. For mutation, we use assignment syntax, but with the location (dynamic or lexical) specified on the left-hand side, e.g. \verb/%x = 1/ or \verb/$x = 1/. In this way we avoid the need to declare variables, and allow ourselves to throw an exception when trying to mutate a non-existent variable, rather than silently creating a new local binding.

\paragraph{Objects.}
For large programs, an object system is desirable as a way to provide user defined types and abstractions. The existing foundations of shell languages already make use of the UNIX \emph{environments} API, which consists of a map of strings to strings, together with a parent pointer. If we allow environments as first-class objects, they can be used in a straightforward manner as prototype-based objects.

A future planned extension would allow an environment to have multiple prototype pointers, with lookup defined in a depth-first manner.

\paragraph{Interruption.}
A core requirement of pipe-based languages is a semantic of process \emph{interruption}: automatic clean-up of processes that will no longer be used. Consider the following Magritte code:
\begin{lstlisting}
  read-lines tmp/large-file
    | take 10
    | each (?line => do-expensive-work <@\$@>line)
\end{lstlisting}

In this example, three processes are spawned: (1) a process with a file open that writes one line at a time to its output, (2) a process that reads 10 times from its input, writes each entry to the output, and then exits, and (3) a process that reads every input and calls a function to perform an expensive task.
A user's intent when typing such code may be to read 10 lines from a file and synchronously perform an action on each line.

In a pipe-based language, we expect all three to be running concurrently, communicating through anonymous pipes or channels.
A user will also expect that, after the first 10 lines are processed, the file is closed, and that all three processes exit.
This expectation is despite the fact that process (1) is specified to read the entire file, and process (3) is a theoretically infinite loop.

In a naive implementation using synchronous channels, process (1) will never be able to write more than 10 lines, and will remain blocked on its output with the file open forever.
Similarly, the \verb/each/ function will never be notified that its input has finished, and will block forever on its input stream.

\paragraph{Compensation.} In order to observe interruption, user code must be able to react to and handle interruption. The most direct way to observe interruption is through a system of \emph{compensation actions}, which are expressions which will be evaluated in the case of an interruption.

For example, the function \verb/file-lines/ above could be implemented as:
\begin{lstlisting}
(file-lines ?fname) = (
  handle = (open-file $fname) %%! close-file $fname
  until (=> eof? $handle) (=> read-until "\n" $handle)
)
\end{lstlisting}

In this code, the \verb/%%!/ symbol registers a compensation action that will be run either at the next checkpoint (the end of the function), or when the process is interrupted. In this way, we can ensure that the file is closed when the process exits. Compensations are the only way in which interruptions can be directly observed by client code---without compensations, the only way of observing forever-blocked processes would be to inspect the process table, or observe the memory footprint of the program. Thus, we can define the primary task of our system as \emph{running a process's compensation actions at the appropriate time}.

\paragraph{Interrupting on Operations vs. Eager Interruption.}

In designing an interruption semantic, we must first decide whether processes are eagerly interrupted when their channels close, or only interrupted on channel operations.

Consider the following example, with the assumption that \\ \verb/do-other-operation/ does not write to the standard output:

\begin{verbatim}
count-forever | ( take 5; do-other-operation ) | take 5
\end{verbatim}

The middle process will write five times, and then continue to do other processing in-thread that does not write to the standard output.
The final \verb/take 5/ operation will return after 5 inputs are read.
It is important to decide, then, whether the middle process should be \emph{interrupted} in the middle of \verb/do-other-operation/, or should it be left alone?

An advantage of lazy interruption is that a user can identify precisely which points in the code have the potential to be interrupted---those points which run a \verb/put/ or a \verb/get/. Whereas with eager-interruption semantics any point in the code may be interrupted, introducing the need for users to either mark critical uninterruptible sections or be very careful about setting correct compensations in the case of mutative algorithms.

On the other hand, lazy interruption has the disadvantage that a producing process must do enough work to produce one more value than will be consumed. If each value is relatively cheap to produce, this is not a problem, but in the case that the values are expensive to produce, this would result in doing a large amount of unnecessary work.

Overall, we have decided that the predictability of lazy interruption is worth the tradeoff for extra values, and we discuss some features that may mitigate this problem in \emph{Future Work}.

\paragraph{Non-closing of Channels.}

Given that multiple processes may be reading from and writing to a channel, it is often the case that a communicating process will end when there are other processes still communicating over the channel. It would not be appropriate in this case to interrupt other processes attached to the channel, as they are still able to communicate.

We define our guiding principle for the appropriate time to interrupt a process as: \emph{A process is interrupted exactly when it can no longer be woken up}.
When a process is blocked on a synchronous channel, it will be woken up as soon as another process communicates on the other end.
Therefore the appropriate time for it to be interrupted is if it is blocked on a channel that will never receive any more operations.
In traditional channel-based systems, this is difficult to detect, because a reference to a channel could be passed anywhere in the program as a standard value, and be interacted with in any way at any time.

Luckily, in pipe-based languages, the arrangement of pipes is usually specified at process spawn time. This enables us to maintain a process register inside of each channel, so that we can decide when all readers or all writers have detached from a channel, at which time the channel closes and interrupts all blocked processes.

\paragraph{Closing Permanence.}
Though we close a channel when all processes on the read or write sides close, there is always the possibility of new processes spawning and attempting to connect to a channel. In such a case, some systems leave open the possibility of \emph{reopening} a channel for further use. In UNIX, for example, the named-pipes API allows for persistent channels (TODO: cite!) that, being merely files, can be reopened at any time.

However, this can lead to some unexpected races between a channel closing and a new process spawning. Consider the example:
\begin{lstlisting}
c = (make-channel)
& count-forever > $c
& take 10 < $c
put 10 > $c
\end{lstlisting}

This example, while contrived, represents an unavoidable race with first-class channels---whether the process running \verb/take 10/ closes the channel before or
after the \verb/put 10/ invocation opens the channel for writing. If we allow channel reopening, we will either block forever, or insert the number 10 into the stream, depending on which happens first. However if we do not allow channel reopening, we can say that, if a process initiates a read or write on a closed channel, it is immediately interrupted. In this case, both sides of the former race have the same termination behavior---the process is closed when \verb/take 10/ returns.

\paragraph{Masking Interruptions.}

A user will often wish to continue after a particular output channel has closed, and in this case will need a facility to mask exceptions for particular channels. Consider the following example, which uses reservoir-sampling to select a uniform random element from a stream:
\begin{lstlisting}
(sample) = (
  hold = (get)
  i = 1
  each (?v =>
    $i = (inc $i)
    # mutate the $hold variable with probability 1/$i
    prob $i 1 && ($hold = %v)
  )

  put $hold
)
\end{lstlisting}

In this example, the function \verb/each/ will consume the entire input stream and mutate a variable with a given probability. After the entire input stream is consumed, we wish to output the resulting value stored in the \verb/$hold/ variable.

However, with the semantics described above, the \verb/each/ function will loop until it receives an interrupt signal from the input channel closing, which will interrupt the entire process and not continue to the following line. In order to avoid this situation, the \verb/each/ function must somehow indicate that it intends to consume the entire stream, and handle the interrupt signal.

While we plan to explore more general mechanisms for error handling, for cases such a this it suffices to provide two builtins, \verb/produce/ and \verb/consume/, each of which takes a single function to loop forever until the standard output or standard input respectively is consumed, whereby control flow continues after the invocation. Using these functions, we might define \verb/each/ as:
\begin{lstlisting}
(each ?fn) = (consume (=> %fn (get)))
\end{lstlisting}

With this definition, the call to \verb/consume/ will mask the interruption, and control flow within the \verb/sample/ function will proceed as normal.

\section{Implementation of Magritte}

The current version of Magritte is implemented as a straightforward interpreter written in Ruby, using Ruby's builtin threads, exceptions, and mutexes to implement processes and channels.

\paragraph{Channel Registry.}

Our channel implementation is a standard implementation of synchronous channels, with the addition of four intrinsic methods, used only internally by the interpreter: \verb/add_reader/, \verb/remove_reader/, \verb/add_writer/, and \verb/remove_writer/, which register and deregister processes as described above. When \verb/remove_reader/ or \verb/remove_writer/ result in an empty registered set, they additionally close the channel and raise an internal \verb/Interrupt/ exception in every blocked thread.

Once the channel is closed, every call to \verb/read/ and \verb/write/ will cause the calling process to be interrupted, as described above in \emph{Closing Permanence}.

In order to reduce unnecessary use of Ruby threads, we also find it is simpler to, instead of registering \emph{processes} to the channels, to register \emph{frames}. In this way, we can register all inputs and outputs on frame entry, and use standard Ruby exception handling to ensure we properly unregister all inputs and outputs on frame exit. This means that different frames in the same process can be redirected to different channels, which makes the \verb/>/ and \verb/</ redirection syntax straightforward to implement---we simply push a new frame with different channels attached.

Cascading of interruptions then happens naturally---when a channel closes, a process is interrupted, causing it to unwind its stack and deregister channels, thereby potentially causing other channels to close.

\paragraph{Spawning Order Dependency.}

It is necessary to take some care with the implementation of the \verb/spawn/ primitive, that we wait until the spawned process has finished registering its channels before the spawning process continues. Consider the example, which outputs 10 numbers, possibly out of order:
\begin{lstlisting}
(drain) = (each (?x => put %x))
count-forever | (& drain; & drain) | take 10
\end{lstlisting}
The middle process is responsible for spawning two processes that funnel data from their input to their output. However, since they are both spawned in the background, the spawning process will immediately return. In general this is not a problem, since the \verb/drain/ processes should be keeping the two pipes open. However, if we do not take care to wait until they have finished registering their channels, there is a risk that the spawning process will return first and close the two pipes.

For this reason, we have special case handling for the root frame---the channels are registered in the \emph{spawning} process, before the spawned process is allowed to start.

\paragraph{Collectors.}
In order to implement the call semantics described above, we also implement a write-only channel called a \emph{collector}. Collectors cannot be created directly by users, but only appear in the interpreter when we evaluate a \verb/Subst/ node, which represents parentheses in argument position.

Naively, collectors would ignore registration commands and simply append written elements to an array. However, it is still necessary to track registered writers. Consider the example:
\begin{lstlisting}
(range ?n) = (count-forever | take %n)
my-list = [(range 10 | (& drain; & drain))]
\end{lstlisting}

In this example, there is an open question of how long we should wait until reading the collection and continuing. If we naively wait until the base command is finished, we will continue early and miss values that may be written later. Thus substitution waits until all writers to the collector have deregistered.

To implement this, we add a \verb/wait_for_close/ intrinsic to collectors, which returns immediately if closed, and otherwise adds the current thread to a waiting set and sleeps. Upon closing the channel in \verb/remove_writer/, all elements of this waiting set are awoken, and flow continues. Thus the algorithm for substiution is:
\begin{lstlisting}
* Create a new collector $c
* Evaluate the contents with standard output into the collector
* Run $c.wait_for_close
\end{lstlisting}

\section{Related Work}

\paragraph{Iterators.}
(TODO: iterators in other languages?)
PEP 533 Authors(TODO: cite properly) describe an automatic closing system for generators, in which an \verb/__iterclose__/ signal is sent by the consumer back to the producer as soon as a consumer exits early.
But channel-based systems generally allow for multiple concurrent readers and writers---so automatically cleaning up a channel when a single consumer exits would cause processes that are still able to communicate to be terminated early.

\paragraph{Go Language.}

The Go programming language contains a robust channel implementation, which can be used to implement a wide variety of concurrent algorithms. In an untyped language similar to Go (the following implementation is not possible in general in Go without generic types), we could implement something similar to our system by adding input and output channels to every function and heavily currying functions:
\begin{lstlisting}[language=C, morekeywords={fn,channel,nil,go}]
fn pipe(lhs, rhs, inp, out) {
  c = channel()
  go fn () { lhs(inp, c); };
  rhs(c, out)
}

collect = collector()
pipe(curry(range, 10), curry(fan, 2, fn(el, inp, out) { write(el + 1, out); }), nil, collect);

return collect.values()
\end{lstlisting}

However, such a system would still require manual use of input and output channels when a simple pipeline would suffice---and the user must still translate between returning vs. outputting functions and expressions. Additionally, Go does not contain any kind of channel closing semantics or automatic process cleanup. Finally, such a system would be far too heavyweight for a REPL context, and would not make for a viable shell.

Python PEP 533; Go, Erlang; Guile Scheme, TCL;

\paragraph{RC Shell.}
The \verb/rc/ shell \cite{duff} from the Plan 9 operating system from Bell Labs introduces a list data type to shell, and cleans up some of the historical surprising quotation behavior of bash. However its list type is not nestable---nested lists are automatically flattened at creation time. Additionally, these lists cannot be passed through channels.

\paragraph{Es Shell.}
The \verb/es/ shell \cite{haahr} attempts to extend the standard shell model with lambda functions with lexical scoping. These shell functions are marshalled into strings when passed through pipes (whether to shell functions or external programs), but are still written byte-by-byte and thus are subject to interleaving in concurrent cases.

\bibliographystyle{ipsjsort-e}
\bibliography{paper}

\end{document}